---
title: "Reproducible Science Assigment"
author: "Laura Barria"
date: "08/10/2026"
output: 
  html_document:
    df_print: paged
---
I will use the data from a study on kinematics of wing beats in flying swallows. I will analyse the relationship between air speed and wing beat amplitude using linear regression. The data is from a study done by Kristy Park and her colleagues in 2001.

```{r}
#libraries####
library(performance)
library(tidyverse)
library(see)


df_park <- read_csv("Parketal2001WBK-2.csv")
```
check for entry mistakes in the data

```{r}
ggplot(aes(x=Airspeed), data= df_park)+
  geom_histogram()
```

```{r}
ggplot(aes(x=Wingbeat_Amp), data = df_park)+
  geom_histogram()
```
```{r}
df_park |> 
  filter(Bird_ID == 1) |> 
  ggplot(aes(x=Airspeed, y=Wingbeat_Amp)) +
  geom_point()+
  theme_bw()
```
```{r}
df_park |> 
  filter(Bird_ID == 1) |> 
  ggplot(aes(x=Airspeed, y=Wingbeat_Amp)) +
  geom_point()+
  xlim(0,14) +
  ylim(40,130)+
  theme_bw() 
```
only analyze one of the two birds

```{r}
df_bird1 <- df_park |> 
  filter(Bird_ID ==1)
```

regression analysis on Bird ID 1

```{r}
lm_wba_1 <- lm(Wingbeat_Amp ~ Airspeed,
               data= df_bird1)
str(lm_wba_1)
```
check the model

```{r}
check_model(lm_wba_1)
```
check the summary of the model 
```{r}
summary(lm_wba_1)

```
check the confidence interval

```{r}

confint(lm_wba_1)
```
from the analysis done it is shown  that with higher wind velocity, the wingbeat amplification increases (Wing aplitude= b = 5.03 ± 0.306, t=16.44, p= 5.08e-08 )


update the model with only a subset of data

```{r}
lm_wba_2 <- update(lm_wba_1, data=df_bird1[-1,])
summary(lm_wba_1)
summary(lm_wba_2)
```
```{r}
lm_wba_3 <- update(lm_wba_1, data=df_bird1[-11,])
summary(lm_wba_1)
summary(lm_wba_2)
summary(lm_wba_3)
```
predict the response variable over a certain range

```{r}
df_newbird <-
  tibble(Airspeed=seq(min(df_bird1$Airspeed), #minimum of the range
         max(df_bird1$Airspeed), #max of the range,
         length=100)) #this tells R how many values to generate
df_pred <- predict(lm_wba_1,
                   newdata=df_newbird,
                   int ="c")
df_newbirds_pred <- bind_cols(df_newbird, as_tibble(df_pred))
```

plotting with ggplot2

```{r}
ggplot(df_newbirds_pred) +
  (aes(x=Airspeed, y=fit)) +
  geom_line()+
  geom_point(data= df_bird1,aes(x=Airspeed,y=Wingbeat_Amp))+
  geom_ribbon(aes(ymin=lwr,
                  ymax=upr),
              alpha=0.25) +
  labs(x= "Airspeed (m/s)",
       y=expression("Wingbeat Amplitude(°)"))+
  theme_bw()
```

